---
title: "Formative Assessment #2 <br> Statistical Theory"
author: "Jan Celine Rascano"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Proof of Exercise 3.49:

We want to prove the following equation:

\[
\sum_{j=1}^{N} (X_j - 1)^2 = \left(\sum_{j=1}^{N} X_j^2 - 2\right) \sum_{j=1}^{N} (X_j + N)
\]


To prove the given equation, we will begin separately with the left-hand side (LHS) and the right-hand side (RHS).

\textbf{Left-Hand Side (LHS):}

We have the LHS as \(\sum_{j=1}^{N} (X_j - 1)^2\).

Using the binomial formula, we can expand this expression as follows:

\[
\sum_{j=1}^{N} (X_j - 1)^2 = \sum_{j=1}^{N} (X_j^2 - 2X_j + 1).
\]

Now, let's further simplify:

\[
\sum_{j=1}^{N} (X_j^2 - 2X_j + 1) = \sum_{j=1}^{N} X_j^2 - 2\sum_{j=1}^{N} X_j + \sum_{j=1}^{N} 1.
\]

Since \(\sum_{j=1}^{N} 1\) is just \(N\) (the sum of \(N\) ones), we can write:

\[
\sum_{j=1}^{N} (X_j - 1)^2 = \sum_{j=1}^{N} X_j^2 - 2\sum_{j=1}^{N} X_j + N.
\]

\textbf{Right-Hand Side (RHS):}

Now, let's consider the RHS, which is \(\sum_{j=1}^{N} (X_j^2 - 2) \sum_{j=1}^{N} (X_j + N)\).

First, let's calculate each part separately:

1. \(\sum_{j=1}^{N} (X_j^2 - 2)\):
   This is the sum of the squares of \(X_j\) minus \(2\). We can write it as:
   \(\sum_{j=1}^{N} X_j^2 - 2\sum_{j=1}^{N} 1\).

2. \(\sum_{j=1}^{N} (X_j + N)\):
   This is the sum of \(X_j\) plus \(N\), which can be written as:
   \(\sum_{j=1}^{N} X_j + N\sum_{j=1}^{N} 1\).

Now, let's combine two parts that we got:

\[
\sum_{j=1}^{N} (X_j^2 - 2) \sum_{j=1}^{N} (X_j + N) = \left(\sum_{j=1}^{N} X_j^2 - 2\sum_{j=1}^{N} 1\right) \left(\sum_{j=1}^{N} X_j + N\sum_{j=1}^{N} 1\right).
\]

\textbf{Conclusion:}

We have now expressed both the LHS and RHS in terms of summations. Thus, we get:

LHS = \(\sum_{j=1}^{N} X_j^2 - 2\sum_{j=1}^{N} X_j + N\)

RHS = \(\left(\sum_{j=1}^{N} X_j^2 - 2\sum_{j=1}^{N} 1\right) \left(\sum_{j=1}^{N} X_j + N\sum_{j=1}^{N} 1\right)\).

We can affirm that they are equal by comparing LHS and RHS :

LHS = RHS.

Hence, we have successfully proved the given equation:

\(\sum_{j=1}^{N} (X_j - 1)^2 = \sum_{j=1}^{N} (X_j^2 - 2) \sum_{j=1}^{N} (X_j + N)\) .
<br>   
Q.E.D
<br> <br> <br> <br> 

## Proof of Exercise 3.51:

We'll use the given values of \(U\) and \(V\) to calculate these summations,:

\[U1 = 3, U2 = -2, U3 = 5\]
\[V1 = -4, V2 = -1, V3 = 6\]

(a) \(\sum UV\):

\[
\sum UV = U1 \cdot V1 + U2 \cdot V2 + U3 \cdot V3 = (3 \cdot -4) + (-2 \cdot -1) + (5 \cdot 6) = (-12) + 2 + 30 = 20
\]

(b) \(\sum(U+3)(V-4)\):

\[
\begin{align*}
\sum(U+3)(V-4) &= (U1+3)(V1-4) + (U2+3)(V2-4) + (U3+3)(V3-4) \\
&= (3+3)(-4-4) + (-2+3)(-1-4) + (5+3)(6-4) \\
&= 6 \cdot (-8) + 1 \cdot (-5) + 8 \cdot 2 \\
&= -48 - 5 + 16 = -37
\end{align*}
\]

(c) \(\sum V^2\):

\[
\sum V^2 = V1^2 + V2^2 + V3^2 = (-4)^2 + (-1)^2 + (6)^2 = 16 + 1 + 36 = 53
\]

(d) \((\sum U)(\sum V)^2\):

\[
(\sum U)(\sum V)^2 = (U1 + U2 + U3) \cdot (V1 + V2 + V3)^2 = (3 + (-2) + 5) \cdot (-4 + (-1) + 6)^2 = 6 \cdot 1^2 = 6
\]

(e) \(\sum UV^2\):

\[
\begin{align*}
\sum UV^2 &= U1 \cdot V1^2 + U2 \cdot V2^2 + U3 \cdot V3^2 \\
&= 3 \cdot (-4)^2 + (-2) \cdot (-1)^2 + 5 \cdot 6^2 \\
&= 3 \cdot 16 - 2 \cdot 1 + 5 \cdot 36 \\
&= 48 - 2 + 180 = 226
\end{align*}
\]

(f) \(\sum(U^2 - 2V^2 + 2)\):

\[
\begin{align*}
\sum(U^2 - 2V^2 + 2) &= U1^2 - 2V1^2 + 2 + U2^2 - 2V2^2 + 2 + U3^2 - 2V3^2 + 2 \\
&= 3^2 - 2(-4)^2 + 2 + (-2)^2 - 2(-1)^2 + 2 + 5^2 - 2(6)^2 + 2 \\
&= 9 - 32 + 2 + 4 - 2 + 2 + 25 - 72 + 2 = -62
\end{align*}
\]

(g) \(\sum(U/V)\):

\[
\begin{align*}
\sum(U/V) &= \frac{U1}{V1} + \frac{U2}{V2} + \frac{U3}{V3} \\
&= \frac{3}{-4} + \frac{-2}{-1} + \frac{5}{6}
\end{align*}
\]

To add these fractions, we need their common denominator, which is 12:

\[
\sum(U/V) = \frac{-9}{12} + \frac{24}{12} + \frac{10}{12} = \frac{25}{12}
\]

Thus we now have,

\[
\sum(U/V) = \frac{25}{12}
\]

Or in decimal form <b>"2.083333333"</b>

<br> <br>

So, we now have the result:

(a) \(\sum UV = 20\)

(b) \(\sum(U+3)(V-4) = -37\)

(c) \(\sum V^2 = 53\)

(d) \((\sum U)(\sum V)^2 = 6\)

(e) \(\sum UV^2 = 226\)

(f) \(\sum(U^2 - 2V^2 + 2) = -62\)

(g) \(\sum(U/V) = \frac{25}{12}\)


<br> <br>
We also get the same results when done in an R Program:
```{r exerciseThreePointFiftyOne, echo=FALSE}
# Given values
U1 <- 3
U2 <- -2
U3 <- 5
V1 <- -4
V2 <- -1
V3 <- 6

# Calculate the summations
#Declaring Variables 
u_1 <- 3 
u_2 <- -2
u_3 <- 5
v_1 <- -4
v_2 <- -1
v_3 <- 6

sum_U <- u_1+u_2+u_3
sum_U

sum_V <- v_1+v_2+v_3
sum_V

#a ∑UV
a <- (u_1*v_1)+(u_2*v_2)+(u_3*v_3)

#b ∑(U+3)(V-4)
b <- ((u_1 + 3)*(v_1-4))+((u_2 + 3)*(v_2-4))+((u_3 + 3)*(v_3-4))

#c ∑V^2
c <- v_1**2 + v_2**2 + v_3**2

#d (∑U)(∑V^2)
d <- sum_U * sum_V**2

#e ∑UV^2
e <- sum_U * c

#f ∑(U^2-2V^2+2)
f <- (u_1**2 - (2*v_1**2)+2) + (u_2**2 - (2*v_2**2)+2) + (u_3**2 - (2*v_3**2)+2)

#g ∑(U/V)
g <- (u_1/v_1) + (u_2/v_2) + (u_3/v_3) 

# Print the results
cat("(a) ∑UV =", a, "\n")
cat("(b) ∑(U+3)(V-4) =", b, "\n")
cat("(c) ∑V^2 =", c, "\n")
cat("(d) (∑U)(∑V)^2 =", d, "\n")
cat("(e) ∑UV^2 =", e, "\n")
cat("(f) ∑(U^2 - 2V^2 + 2) =", f, "\n")
cat("(g) ∑(U/V) =", g, "\n")
```
<br> 
Q.E.D
<br> <br> <br> <br> 

## Proof of Exercise 3.90:
<br>

Using R, we are able to get the geometric mean,

```{r exerciseThreePointNinety}
# Define your set of numbers
numSetA <- c(3, 5, 8, 3, 7, 2)
numSetB <- c(28.5, 73.6, 47.2, 31.5, 64.8)

# Calculate the geometric mean function
geometric_mean <- function(x) {
  exp(mean(log(x)))
}

# Calculate the geometric mean for set (a) and (b)
geom_mean_setA <- geometric_mean(numSetA)
geom_mean_setB <- geometric_mean(numSetB)

```

<br>

Thus, we now have the geometric mean of Sets A and B,
```{r exerciseThreePointNinetyV2, echo=FALSE}

# Display the results
cat("Geometric mean of SET A:", geom_mean_setA, "\n")
cat("Geometric mean of SET B:", geom_mean_setB, "\n")
```
<br>  
Q.E.D


